### 2024-04-17 paper/manuscript.qmd
add apaquarto doc

---
title: "Is Phenomenological Control related to Illusion Sensitivity"
# If blank, the running header is the title in upper case.
shorttitle: "Illusion Sensitivity and Phenomenological Control"
# Set names and affiliations.
# It is nice to specify everyone's orcid, if possible.
# There can be only one corresponding author, but declaring one is optional.
author:
  - name: Dominique Makowski
    corresponding: true
    orcid: 0000-0001-5375-9967
    email: D.Makowski@sussex.ac.uk
    url: https://realitybending.github.io/
    # Roles are optional. 
    # Select from the CRediT: Contributor Roles Taxonomy https://credit.niso.org/
    # conceptualization, data curation, formal Analysis, funding acquisition, investigation, 
    # methodology, project administration, resources, software, supervision, validation, 
    # visualization, writing, editing
    roles:
      - Conceptualization
      - Data curation
      - formal Analysis
      - Funding acquisition
      - Investigation
      - Methodology
      - Project administration
      - Resources
      - Software
      - Supervision
      - Validation
      - Visualization
      - Writing – original draft
      - Writing – review & editing
    affiliations:
      - id: id1
        name: "University of Sussex"
        # group: Clinical Psychology Program
        department: School of Psychology
        # address: 1234 Capital St.
        # city: Albany
        # region: NY
        # postal-code: 12084-1234
  - name: Ana Neves
    orcid: 0000-0000-0000-0002
    roles:
      - Project administration
      - Data curation
      - Formal Analysis
      - Investigation
      - Visualization
      - Writing – original draft
      - Writing – review & editing
    affiliations: 
      - ref: id1
  - name: Zoltan Dienes
    orcid: 0000-0001-7454-3161
    roles:
      - Supervision
      - Writing – review & editing
    affiliations: 
      - ref: id1
author-note:
  status-changes: 
    # Example: [Author name] is now at [affiliation].
    affiliation-change: ~
    # Example: [Author name] is deceased.
    deceased: ~
  # Disclosures condensed to one paragraph, but you can start a field with two line breaks to break them up: \n\nNew Paragraph
  disclosures:
    # Example: This study was registered at X (Identifier Y).
    study-registration: ~
    # Acknowledge and cite data/materials to be shared.
    data-sharing: ~
    # Example: This article is based on data published in [Reference].
    # Example: This article is based on the dissertation completed by [citation].  
    related-report: ~
    # Example: [Author name] has been a paid consultant for Corporation X, which funded this study.
    conflict-of-interest: ~
    # Example: This study was supported by Grant [Grant Number] from [Funding Source].
    financial-support: ~
    # Example: The authors are grateful to [Person] for [Reason].
    gratitude: ~
    # Example. Because the authors are equal contributors, order of authorship was determined by a fair coin toss.
    authorship-agreements: ~
abstract: "Blablablabla blablabla todo todo todo."
# Put as many keywords at you like, separated by commmas (e.g., [reliability, validity, generalizability])
keywords: [keyword1, keyword2, keyword3]
# If true, tables and figures are mingled with the text instead of listed at the end of the document.
floatsintext: true
# Numbered lines (.pdf and .docx only)
numbered-lines: true
# File with references
bibliography: bibliography.bib
# Suppress title page
suppress-title-page: false
# Masks references that appear in the masked-citations list
mask: false
masked-citations:
  - schneider2012cattell
  - schneider2015intelligence
# Language options. See https://quarto.org/docs/authoring/language.html
lang: en
language:
  citation-last-author-separator: "and"
  citation-masked-author: "Masked Citation"
  citation-masked-date: "n.d."
  citation-masked-title: "Masked Title"
  title-block-author-note: "Author Note"
  title-block-correspondence-note: "Correspondence concerning this article should be addressed to"
  title-block-role-introduction: "Author roles were classified using the Contributor Role Taxonomy (CRediT; https://credit.niso.org/) as follows:"
  references-meta-analysis: "References marked with an asterisk indicate studies included in the meta-analysis."
format:
  apaquarto-docx: default
  apaquarto-html: default
  apaquarto-pdf:
    # Can be jou (journal), man (manuscript), stu (student), or doc (document)
    documentmode: man
---

```{r}
#| label: setup
#| include: false
library(conflicted)
library(tidyverse)
library(flextable)
library(ftExtra)
library(officer)
library(knitr)

conflicts_prefer(dplyr::filter, .quiet = TRUE)
conflicts_prefer(flextable::separate_header, .quiet = TRUE)
```

This is my introductory paragraph. The title will be placed above it automatically. *Do not start with an introductory heading* (e.g., "Introduction"). The title acts as your Level 1 heading for the introduction.

See [here](https://quarto.org/docs/authoring/footnotes-and-citations.html) for instructions on setting up citations and references.

A parenthetical citation requires square brackets [@CameronTrivedi2013]. This reference was in my bibliography file. An in-text citation is done like so:

@CameronTrivedi2013 make some important points ...

See [here](https://wjschne.github.io/apaquarto/writing.html#references) for explanations, examples, and citation features exclusive to apaquarto. For example, apaquarto can automatically handle possessive citations:

# References

<!-- References will auto-populate in the refs div below -->

::: {#refs}
:::