### 2025-04-17 analysis/1_cleaning.qmd
finished data_cleaning script

---
title: "Illusion Game & PCS"
subtitle: "Preprocessing"
date: "`r Sys.Date()`"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: false
    toc: true
execute: 
  cache: true
---

# Data Preparation

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(ggdist)
library(ggside)
library(easystats)
library(patchwork)

data_ppt <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGamePhenomenologicalControl/refs/heads/main/data/rawdata_participants.csv")

data_ig <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGamePhenomenologicalControl/refs/heads/main/data/rawdata_illusion.csv")|> 
  group_by(Illusion_Type) |> 
  mutate(Illusion_Effect = ifelse(sign(Illusion_Strength) > 0, "Incongruent", "Congruent"),
         Task_Difficulty = abs(Illusion_Difference),
         Condition_Illusion = datawizard::categorize(
           Illusion_Strength, split="quantile", n_groups=4,
           labels=c("Congruent - Strong", "Congruent - Mild", "Incongruent - Mild", "Incongruent - Strong")),
         Condition_Difficulty = datawizard::categorize(Task_Difficulty, split="quantile", n_groups=2, labels=c("Hard", "Easy"))) |> 
  ungroup()

```

The initial sample consisted of `r report::report_participants(data_ppt, age="Age", gender="Gender", education="Education")`, for a total trial number of `r nrow(data_ppt)`.

```{r}

```

## Score Computation

### IPIP6

```{r}
SD_per_dim <- function(x, dims="") {
  m <- matrix(nrow=nrow(x), ncol=0)
  for(s in dims) {
    m <- cbind(m, sapply(as.data.frame(t(x[grepl(s, names(x))])), sd))
  }
  m
}
```

```{r}
#| code-fold: false

ipip6 <- select(data_ppt, starts_with("ipip_"))
ipip6[grepl("_R", names(ipip6))] <- 8 - ipip6[grepl("_R", names(ipip6))]

data_ppt$IPIP6_Extraversion <- rowMeans(ipip6[grepl("ext", names(ipip6))])
data_ppt$IPIP6_Conscientiousness <- rowMeans(ipip6[grepl("cons", names(ipip6))])
data_ppt$IPIP6_Neuroticism <- rowMeans(ipip6[grepl("neuro", names(ipip6))])
data_ppt$IPIP6_Openness <- rowMeans(ipip6[grepl("open", names(ipip6))])
data_ppt$IPIP6_HonestyHumility <- rowMeans(ipip6[grepl("honhum", names(ipip6))])
data_ppt$IPIP6_Agreeableness <- rowMeans(ipip6[grepl("agr", names(ipip6))])
data_ppt$IPIP6_SD <- rowMeans(SD_per_dim(ipip6, c("Extraversion", "Conscientiousness", "Neuroticism", "Openness", "HonestyHumility", "Agreeableness")))

# data_ppt$IPIP6_Extraversion <- rowMeans(ipip6[grepl("Ext", names(ipip6))])
# data_ppt$IPIP6_Conscientiousness <- rowMeans(ipip6[grepl("Cons", names(ipip6))])
# data_ppt$IPIP6_Neuroticism <- rowMeans(ipip6[grepl("Neuro", names(ipip6))])
# data_ppt$IPIP6_Openness <- rowMeans(ipip6[grepl("Ppen", names(ipip6))])
# data_ppt$IPIP6_HonestyHumility <- rowMeans(ipip6[grepl("HonHum", names(ipip6))])
# data_ppt$IPIP6_Agreeableness <- rowMeans(ipip6[grepl("Agr", names(ipip6))])
# data_ppt$IPIP6_SD <- rowMeans(SD_per_dim(ipip6, c("Extraversion", "Conscientiousness", "Neuroticism", "Openness", "HonestyHumility", "Agreeableness")))
```

### PID-5

```{r}
#| code-fold: false

pid5 <- select(data_ppt, starts_with("pid_"))

data_ppt$PID5_Disinhibition <- rowMeans(pid5[grepl("Dis", names(pid5))])
data_ppt$PID5_Detachment <- rowMeans(pid5[grepl("Det", names(pid5))])
data_ppt$PID5_NegativeAffect <- rowMeans(pid5[grepl("NegAff", names(pid5))])
data_ppt$PID5_Antagonism <- rowMeans(pid5[grepl("Ant", names(pid5))])
data_ppt$PID5_Psychoticism <- rowMeans(pid5[grepl("Psy", names(pid5))])
data_ppt$PID5_SD <- rowMeans(SD_per_dim(pid5, c("Disinhibition", "Detachment", "NegativeAffect",
                                             "Antagonism", "Psychoticism")))
```

### Experiment Duration

The experiment's median duration is `r insight::format_value(median(data_ppt$Experiment_Duration))` min (`r insight::format_ci(bayestestR::hdi(data_ppt$Experiment_Duration, ci=0.5))`).

```{r}
data_ppt |>
  mutate(Participant = fct_reorder(Participant, Experiment_Duration),
         Category = ifelse(Experiment_Duration > 50, "extra", "ok"),
         Duration = ifelse(Experiment_Duration > 50, 50, Experiment_Duration)) |>
  ggplot(aes(y = Participant, x = Duration)) +
  geom_point(aes(color = Category, shape = Category)) +
  geom_vline(xintercept = median(data_ppt$Experiment_Duration), color = "red", linetype = "dashed") +
  scale_shape_manual(values = c("extra" = 3, ok = 19)) +
  scale_color_manual(values = c("extra" = "red", ok = "black")) +
  guides(color = "none", shape = "none") +
  ggside::geom_xsidedensity(fill = "grey", color=NA) +
  ggside::scale_xsidey_continuous(expand = c(0, 0)) +
  labs(
    title = "Experiment Completion Time",
    x = "Duration (in minutes)",
    y = "Participants"
  )  +
  see::theme_modern() +
  ggside::theme_ggside_void() +
  theme(ggside.panel.scale = .3, 
        axis.text.y = element_blank()) 
```

# Exclusion

```{r}
#| echo: true
#| code-fold: false

outliers <- list()
outliers_half <- list()
```

## Reaction Time (per Block)

```{r}
#| message: false

errorrate <- data_ig |>
  group_by(Participant, Illusion_Type, Block) |>
  summarize(ErrorRate_per_block = sum(Error) / n()) |>
  ungroup() |>
  arrange(desc(ErrorRate_per_block))

d_all <- estimate_density(filter(data_ig, RT < 3)$RT)

plot_distribution <- function(dat) {

  data_error <- errorrate |>
    filter(Participant %in% unique(dat$Participant)) |>
    group_by(Participant, Block) |>
    summarize(y = mean(ErrorRate_per_block), .groups="drop") |>
    mutate(x = ifelse(Block == "A", 2.1, 2.3),
           color = case_when(
              Participant %in% outliers ~ "red",
              Participant %in% outliers_half ~ "orange",
              TRUE ~ "blue"
            ))

  dat |>
    filter(RT < 3) |>
    estimate_density(select = "RT", at = c("Participant", "Block")) |>
    group_by(Participant) |>
    normalize(select = "y") |>
    ungroup() |>
    mutate(
      Participant = fct_relevel(Participant, sort(unique(dat$Participant))),
      color = case_when(
        Participant %in% outliers ~ "red",
        Participant %in% outliers_half ~ "orange",
        TRUE ~ "blue"
      )
    ) |>
    ggplot(aes(x = x, y = y)) +
    geom_bar(data = data_error, aes(fill = color), stat = "identity", width=0.19) +
    geom_segment(aes(x = 2, xend = 2.4, y = 0.5, yend = 0.5), color = "black", linetype="dashed", linewidth = 0.5) +
    geom_area(data = normalize(d_all, select = "y"), alpha = 0.2) +
    geom_line(aes(color = color, group = interaction(Participant, Block), linetype = Block), linewidth = 0.8) +
    # geom_vline(xintercept = 0.125, linetype = "dashed", color = "red", size = 0.5) +
    scale_color_manual(values = c("red" = "#F44336", "orange" = "#FF9800", "blue" = "blue"), guide = "none") +
    scale_fill_manual(values = c("red" = "#F44336", "orange" = "#FF9800", "blue" = "blue"), guide = "none") +
    scale_x_continuous(expand = c(0, 0), breaks=c(0, 0.5, 1, 1.5, 2), labels=c("0", "0.5", "1", "1.5", "2")) +
    scale_y_continuous(expand = c(0, 0)) +
    coord_cartesian(xlim = c(0, 2.4)) +
    theme_modern() +
    theme(axis.text.y = element_blank(),
          axis.line.y = element_blank(),
          strip.text.x = element_text(size = rel(1.5)),
          legend.position = "none") +
    facet_wrap(~Participant, nrow=10) +
    labs(y = "", x = "Reaction Time (s)")
}
```

### 1-100

```{r}
#| fig-width: 15
#| fig-height: 15
#| eval: false

plot_distribution(data_ig[data_ig$Participant %in% data_ppt$Participant[1:100],])
```

### 101-200

```{r}
#| fig-width: 15
#| fig-height: 15
#| eval: false

plot_distribution(data_ig[data_ig$Participant %in% data_ppt$Participant[1:200],])

```

:::

We discarded `r length(outliers)` participants (entirely) and `r length(outliers_half)` participant's second blocks.

```{r}
# data_ig <- data_ig |> 
#   filter(!Participant %in% outliers) |> 
#   filter(!((Block == "B") & (Participant %in% outliers_half)))
```

## Error Rate (per Block)

For each block, we computed the error rate and, if more than 50%, we discarded the whole block (as it likely indicates that instructions got mixed up, for instance participants were selecting the smaller instead of the bigger circle).

```{r p_outliers_errorblock}
#| eval: false

errorrate |>
  estimate_density(at = c("Illusion_Type", "Block"), method = "KernSmooth") |>
  ggplot(aes(x = x, y = y)) +
  geom_line(aes(color = Illusion_Type, linetype = Block)) +
  geom_vline(xintercept = 0.5, linetype = "dashed") +
  scale_x_continuous(expand = c(0, 0), labels = scales::percent) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_color_manual(values = c("Ebbinghaus" = "#2196F3", "MullerLyer" = "#4CAF50", "VerticalHorizontal" = "#FF5722")) +
  labs(y = "Distribution", x = "Error Rate") +
  theme_modern()

remove_badblocks <- function(df) {
  n <- nrow(df)
  df <- df |>
    group_by(Participant, Illusion_Type, Block) |>
    mutate(ErrorRate_per_block = sum(Error) / n()) |>
    ungroup() |>
    filter(ErrorRate_per_block < 0.5) |>
    select(-ErrorRate_per_block)

  text <- paste0(
    "We removed ",
    n - nrow(df),
    " (",
    insight::format_value((n - nrow(df)) / n, as_percent = TRUE),
    ") trials belonging to bad blocks."
  )
  list(data = df, text = text)
}

out <- remove_badblocks(data_ig)
print(paste("Illusion (session 1):", out$text))
data_ig <- out$data
```

## Reaction Time (per Block)

```{r outliers_RTtrials}
#| eval: false

check_trials <- function(df) {
  data <- df |>
    mutate(Outlier = ifelse(RT >= 10, TRUE, FALSE)) |>
    group_by(Participant) |>
    mutate(Outlier = ifelse(RT < 0.150 | standardize(RT, robust = TRUE) > 4, TRUE, Outlier)) |>
    ungroup()

  p1 <- data |>
    filter(RT < 10) |>
    estimate_density(select = "RT", at = "Participant") |>
    group_by(Participant) |>
    normalize(select = "y") |>
    ungroup() |>
    merge(data |>
      group_by(Participant) |>
      mutate(Threshold = median(RT) + 4 * mad(RT)) |>
      filter(Error == 0) |>
      summarize(Threshold = mean(Threshold))) |>
    mutate(Outlier = ifelse(x >= Threshold, TRUE, FALSE)) |>
    ggplot(aes(x = x, y = y)) +
    geom_area(data = normalize(estimate_density(filter(data, RT < 10), select = "RT"), select = "y"), alpha = 0.2) +
    geom_line(aes(color = Participant, linetype = Outlier), alpha = 0.2) +
    geom_vline(xintercept = c(125), linetype = "dashed", color = "red") +
    scale_color_material_d("rainbow", guide = "none") +
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0)) +
    guides(linetype = "none") +
    coord_cartesian(xlim = c(0, 5)) +
    theme_modern() +
    theme(axis.text.y = element_blank()) +
    labs(y = "", x = "Reaction Time (s)")

  p2 <- data |>
    group_by(Participant) |>
    summarize(Outlier = sum(Outlier) / n()) |>
    mutate(Participant = fct_reorder(Participant, Outlier)) |>
    ggplot(aes(x = Participant, y = Outlier)) +
    geom_bar(stat = "identity", aes(fill = Participant)) +
    scale_fill_material_d("rainbow", guide = "none") +
    scale_x_discrete(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0), labels = scales::percent) +
    see::theme_modern() +
    theme(axis.text.x = element_blank()) +
    labs(y = "Percentage of outlier trials")

  text <- paste0(
    "We removed ",
    sum(data$Outlier),
    " (",
    insight::format_value(sum(data$Outlier) / nrow(data), as_percent = TRUE),
    ") outlier trials (150 ms < RT < 4 MAD above median)."
  )

  data <- filter(data, Outlier == FALSE)
  data$Outlier <- NULL

  list(p = p1 / p2, data = data, text = text)
}
```

```{r p_outliers_RTtrials}
#| results: 'asis'
#| eval: false

out <- check_trials(data_ig)
out$text
out$p
data_ig <- out$data
```

## Manipulation/Attention Checks

```{r}

outliers$Hello <- data_ppt |>
  dplyr::filter(pcs_hello != "Hello")

outliers$Hello <- outliers$Hello$Participant

```

We removed `r length(outliers$Hello)` (`r insight::format_percent(length(outliers$Hello) / nrow(data_ppt))`) participants that did not reported hearing "Hello" at the beggining of the experiment.

```{r}
# remove after more pilots
outliers$Balls <- data_ppt |>
  dplyr::filter(pcs_noballs == TRUE) # remove after more pilots

# outliers$balls <- data_ppt |>
#   dplyr::filter(pcs_noballs == "TRUE")

outliers$Balls <- outliers$Balls$Participant

```

We removed `r length(outliers$Balls)` (`r insight::format_percent(length(outliers$Balls) / nrow(data_ppt))`) participants that responded that no balls were presented in the screen.

```{r}

outliers$Press <- data_ppt |>
  dplyr::filter(pcs_press < 5)

outliers$Press <- outliers$Press$Participant
```

We removed `r length(outliers$Press)` (`r insight::format_percent(length(outliers$Press) / nrow(data_ppt))`) participants that responded that no balls were presented in the screen.

```{r}
data_ppt <- filter(data_ppt, !Participant %in% outliers$Hello)
data_ppt <- filter(data_ppt, !Participant %in% outliers$Balls)
data_ppt <- filter(data_ppt, !Participant %in% outliers$Press)

data_ig <- filter(data_ig, !Participant %in% outliers$Hello)
data_ig <- filter(data_ig, !Participant %in% outliers$Balls)
data_ig <- filter(data_ig, !Participant %in% outliers$Press)
```

# Final Sample

```{r}
# data_ppt<- filter(data_ppt, !Participant %in% outliers)
# data_ig <- filter(data_ig, Participant %in% data_ppt$Participant)
```

### Age

```{r}
p_age <- estimate_density(data_ppt$Age) |>
  normalize(select = y) |>
  mutate(y = y * 86) |>  # To match the binwidth
  ggplot(aes(x = x)) +
  geom_histogram(data=dfsub, aes(x = Age, fill=Gender), bins=28) +
  # geom_line(aes(y = y), color = "orange", linewidth=2) +
  geom_vline(xintercept = mean(dfsub$Age), color = "red", linewidth=1.5) +
  # geom_label(data = data.frame(x = mean(df$Age) * 1.15, y = 0.95 * 75), aes(y = y), color = "red", label = paste0("Mean = ", format_value(mean(df$Age)))) +
  scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange")) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Age", y = "Number of Participants", color = NULL, subtitle = "Distribution of participants' age") +
  theme_modern(axis.title.space = 10) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(1.1)),
    axis.title.x = element_blank()
  )
p_age
```

### Education

```{r}
p_edu <- data_ppt |>
  mutate(Education = fct_relevel(Education, "Other", "High School", "Bachelor", "Master", "Doctorate")) |>
  ggplot(aes(x = Education)) +
  geom_bar(aes(fill = Education)) +
  scale_y_continuous(expand = c(0, 0), breaks= scales::pretty_breaks()) +
  scale_fill_viridis_d(guide = "none") +
  labs(title = "Education", y = "Number of Participants", subtitle = "Participants per achieved education level") +
  theme_modern(axis.title.space = 15) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(1.1)),
    axis.title.x = element_blank()
  )
p_edu
```

### Ethnicity

```{r p_ethnicity}
ggwaffle::waffle_iron(data_ppt, ggwaffle::aes_d(group = Ethnicity), rows=10) |> 
  ggplot(aes(x, y, fill = group)) + 
  ggwaffle::geom_waffle() + 
  coord_equal() + 
  scale_fill_flat_d() + 
  ggwaffle::theme_waffle() +
  labs(title = "Self-reported Ethnicity", subtitle = "Each square represents a participant", fill="")  +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2)),
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
  )
```

### Summary

The final sample includes `r report::report_participants(dfsub, age="Age", sex="Sex", education="Education")`.

## Save

```{r}
#| code-fold: false

write.csv(data_ppt, "../data/data_participants.csv", row.names = FALSE)
write.csv(data_ig, "../data/data_IllusionGame.csv", row.names = FALSE)
```