### 2025-04-30 analysis/2_analysis.qmd

---
title: "Illusion Game & PCS"
subtitle: "Analysis"
date: "`r Sys.Date()`"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: false
    toc: true
    warning: false 
    message: false 
execute: 
  cache: true
---

## Data Preparation

```{r}
library(tidyverse)
library(ggdist)
library(ggside)
library(easystats)
library(patchwork)
library(psych)

# data_ppt <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGamePhenomenologicalControl/refs/heads/main/data/rawdata_participants.csv")

data_ppt <- read.csv("../data/data_participants.csv")

data_ig <- read.csv("../data/data_IllusionGame.csv") |>
  group_by(Illusion_Type) |> 
  mutate(Illusion_Side = sign(Illusion_Difference), 
    Illusion_Effect = ifelse(sign(Illusion_Strength) > 0, "Incongruent", "Congruent"),
    Task_Difficulty = abs(Illusion_Difference),
      Illusion_Group = case_when(
        Illusion_Effect == "Congruent" ~ "Congruent",
        Illusion_Effect == "Incongruent" & Illusion_Strength <= median(Illusion_Strength[Illusion_Effect == "Incongruent"], na.rm = TRUE) ~ "Mild",
        Illusion_Effect == "Incongruent" & Illusion_Strength > median(Illusion_Strength[Illusion_Effect == "Incongruent"], na.rm = TRUE) ~ "Strong"),
    Condition_Difficulty = datawizard::categorize(Task_Difficulty, split="quantile", n_groups=2, labels=c("Hard", "Easy"))) |> 
  ungroup() |>
  mutate(
    Block = as.factor(Block),
    Illusion_Side = as.factor(Illusion_Side),
    Illusion_Effect = fct_relevel(as.factor(Illusion_Effect), "Incongruent", "Congruent"),
    Response_Correct = as.factor(Response_Correct)
  )

```

## PCS

### Compute subjective scale score

**Note:** Mean subjective scale responses for item 4 (Taste) and item 10 (Post-session suggestion) are each generated from the combination of two responses which require processing prior to generating an overall subjective score. 
The subjective score for item 4 is the mean of the responses to the (a)‘sour’ and (b)‘sweet’ sub-scales for the taste suggestion. 
The subjective score for item 10 is the geometric mean (the square root of the product) of he responses to the (a) ‘urge’ and (b) ‘amnesia’ sub-scales for the post-session suggestion.

```{r}
data_ppt$pcs_tastescore <- rowMeans(data_ppt[grepl("taste", names(data_ppt))])
data_ppt$pcs_postsessionsuggetsion <- sqrt(data_ppt$pcs_urgepress * data_ppt$pcs_memorypress)
data_ppt$pcs <- rowMeans(data_ppt[, c("pcs_handslow", "pcs_mosquito", "pcs_magnetichands", "pcs_armrigidity", "pcs_tastescore", "pcs_postsessionsuggetsion", "pcs_armimmobile", "pcs_music", "pcs_negativehallucination", "pcs_amnesia")], na.rm = TRUE)
data_ppt$pcs_SD <- sd(data_ppt$pcs, na.rm = TRUE)

```

### Reliability

```{r}

pcs_rel <- data_ppt |>
  select("pcs_handslow", "pcs_mosquito", "pcs_magnetichands", "pcs_armrigidity", "pcs_tastescore", "pcs_postsessionsuggetsion", "pcs_armimmobile", "pcs_music", "pcs_negativehallucination", "pcs_amnesia") |>
  psych::alpha()

```

## Illusion Sensitivity Scores

### Error Rate

```{r}
errorrate <- data_ig |>
  group_by(Participant, Illusion_Type, Illusion_Group) |>
  summarize(ErrorRate = sum(Error) / n()) |>
  ungroup() |>
  as.data.frame()
```

### IES

**Note.** The IES is calculated by dividing the *average correct* RT by the *proportion of correct responses*, for each illusion and for each illusion strenght group: Strong Illusion Strength & Incongruent, Mild Illusion Strength & Incongruent, and Congruent

```{r}

IES <- data_ig |>
  group_by(Participant, Illusion_Type, Illusion_Group) |>
  summarise(IES = mean(RT[Response_Correct == "true"] / mean(Response_Correct == "true"))) |>
  ungroup() |>
  as.data.frame()
```

```{r}

data_vs <- left_join(IES, errorrate, by= c("Participant", "Illusion_Type", "Illusion_Group"))
```

### Correlations between Strong and Mild

**Note**.  If correlations high between mild and strong conditions then the mild and strong illusion strength groups will be collapsed and the outcomes will be recomputed across all trials, otherwise they will be treated as separate in subsequent analyses.

```{r}

compute_correlation <- function(data, illusion_type) {
  illusion_data <- data |> dplyr::filter(Illusion_Group != "Congruent", Illusion_Type == illusion_type) |>
    pivot_wider(id_cols = Participant,
      names_from = Illusion_Group,
      values_from = c(ErrorRate, IES),
      names_sep = "_")

  ER_Incongruent <- BayesFactor::correlationBF(
    y = illusion_data$ErrorRate_Mild, 
    x = illusion_data$ErrorRate_Strong
  )

  IES_Incongruent <- BayesFactor::correlationBF(
    y = illusion_data$IES_Mild, 
    x = illusion_data$IES_Strong
  )

  return(list(
    ER_Incongruent = ER_Incongruent,
    IES_Incongruent = IES_Incongruent
  ))
}

```

```{r}

cor_ebbinghaus <- compute_correlation(data_vs, "Ebbinghaus")
cor_mullerlyer <- compute_correlation(data_vs, "MullerLyer")
cor_verticalhorizontal <- compute_correlation(data_vs, "VerticalHorizontal")
```

```{r}

er_ebbinghaus_post<- describe_posterior(cor_ebbinghaus$ER_Incongruent)
er_mullerlyer_post <- describe_posterior(cor_mullerlyer$ER_Incongruent)
er_verticalhorizontal_post <- describe_posterior(cor_verticalhorizontal$ER_Incongruent)
```

```{r}

ies_ebbinghaus_post <- describe_posterior(cor_ebbinghaus$IES_Incongruent)
ies_mullerlyer_post <- describe_posterior(cor_mullerlyer$IES_Incongruent)
ies_verticalhorizontal_post <- describe_posterior(cor_verticalhorizontal$IES_Incongruent)

```

### Colapsed scores

```{r}

# collapsed_groups_data <- data_vs |>
#  filter(Illusion_Group %in% c("Mild", "Strong")) |>
#   group_by(Participant, Illusion_Type) |>
#   summarise(
#     Illusion_Group = "Incongruent",
#     ErrorRate = mean(ErrorRate, na.rm = TRUE),
#     IES = mean(IES, na.rm = TRUE),
#     .groups = "drop"
#   )

# data_vs <- bind_rows(data_vs, collapsed_groups_data) |>
#   arrange(Participant, Illusion_Type, factor(Illusion_Group, levels = c("Mild", "Strong", "Congruent", "Incongruent"))) |>
#   filter(Illusion_Group == c( "Congruent", "Incongruent"))

```

```{r}
data_vs_wide <- data_vs|>
  select(Participant, Illusion_Type, Illusion_Group, ErrorRate, IES) |>
  unite("group", Illusion_Type, Illusion_Group, sep = "_") |>
  pivot_wider(names_from = group, values_from = c(ErrorRate, IES)) |>
  as.data.frame()
```

### Illusion Manipulation Checks

**Note:** Significant differences between the Congruent and Mild and the Mild and Strong groups will be taken as evidence that illsuions functioned as intended.

```{r}
compute_ttest <- function(data, illusion_type) {
  illusion_data <- data |> dplyr::filter(Illusion_Type == illusion_type)

  ERcongruent_vs_mild <- BayesFactor::ttestBF(
    x = illusion_data$ErrorRate[illusion_data$Illusion_Group == "Congruent"],
    y = illusion_data$ErrorRate[illusion_data$Illusion_Group == "Mild"],
    paired = TRUE
  )

  IEScongruent_vs_mild <- BayesFactor::ttestBF(
    x = illusion_data$IES[illusion_data$Illusion_Group == "Congruent"],
    y = illusion_data$IES[illusion_data$Illusion_Group == "Mild"],
    paired = TRUE
  )

  ERmild_vs_strong <- BayesFactor::ttestBF(
    x = illusion_data$ErrorRate[illusion_data$Illusion_Group == "Mild"],
    y = illusion_data$ErrorRate[illusion_data$Illusion_Group == "Strong"],
    paired = TRUE
  )

  IESmild_vs_strong <- BayesFactor::ttestBF(
    x = illusion_data$IES[illusion_data$Illusion_Group == "Mild"],
    y = illusion_data$IES[illusion_data$Illusion_Group == "Strong"],
    paired = TRUE
  )

  # Congruent vs Incongruent if correlation high between mild and strong

  # ERcong_vs_incon < BayesFactor::ttestBF(
  #   x = illusion_data$ErrorRate[illusion_data$Illusion_Group == "Congruent"],
  #   y = illusion_data$ErrorRate[illusion_data$Illusion_Group == "Incongruent"],
  #   paired = TRUE
  # )
  # 
  # IEScong_vs_incon <- BayesFactor::ttestBF(
  #   x = illusion_data$IES[illusion_data$Illusion_Group == "Congruent"],
  #   y = illusion_data$IES[illusion_data$Illusion_Group == "Incongruent"],
  #   paired = TRUE
  # )

  return(list(
    ERcongruent_vs_mild = ERcongruent_vs_mild,
    ERmild_vs_strong = ERmild_vs_strong,
    IEScongruent_vs_mild = IEScongruent_vs_mild,
    IESmild_vs_strong = IESmild_vs_strong
    # ERcong_vs_incon = ERcong_vs_incon,
    # IEScong_vs_incon = IEScong_vs_incon
  ))
}

```

```{r}

ttest_ebbinghaus <- compute_ttest(data_vs, "Ebbinghaus")
ttest_mullerlyer <- compute_ttest(data_vs, "MullerLyer")
ttest_verticalhorizontal <- compute_ttest(data_vs, "VerticalHorizontal")
```

```{r}

er_ebbinghaus_post_cm <- describe_posterior(ttest_ebbinghaus$ERcongruent_vs_mild)
er_ebbinghaus_post_ms <- describe_posterior(ttest_ebbinghaus$ERmild_vs_strong)
# er_ebbinghaus_post_ci <- describe_posterior(ttest_ebbinghaus$ERcong_vs_incon)

er_mullerlyer_post_cm <- describe_posterior(ttest_mullerlyer$ERcongruent_vs_mild)
er_mullerlyer_post_ms <- describe_posterior(ttest_mullerlyer$ERmild_vs_strong)
# er_mullerlyer_post_ci <- describe_posterior(ttest_mullerlyer$ERcong_vs_incon)

er_verticalhorizontal_post_cm <- describe_posterior(ttest_verticalhorizontal$ERcongruent_vs_mild)
er_verticalhorizontal_post_ms <- describe_posterior(ttest_verticalhorizontal$ERmild_vs_strong)
# er_verticalhorizontal_post_ci <- describe_posterior(ttestcor_verticalhorizontal$ERcong_vs_incon)

```

```{r}

IES_ebbinghaus_post_cm <- describe_posterior(ttest_ebbinghaus$IEScongruent_vs_mild)
IES_ebbinghaus_post_ms <- describe_posterior(ttest_ebbinghaus$IESmild_vs_strong)
# IES_ebbinghaus_post_ci <- describe_posterior(ttest_ebbinghaus$IEScong_vs_incon)

IES_mullerlyer_post_cm <- describe_posterior(ttest_mullerlyer$IEScongruent_vs_mild)
IES_mullerlyer_post_ms <- describe_posterior(ttest_mullerlyer$IESmild_vs_strong)
# IES_mullerlyer_post_ci <- describe_posterior(ttest_ebbinghaus$IEScong_vs_incon)

IES_verticalhorizontal_post_cm <- describe_posterior(ttest_verticalhorizontal$IEScongruent_vs_mild)
IES_verticalhorizontal_post_ms <- describe_posterior(ttest_verticalhorizontal$IESmild_vs_strong)
# IES_verticalhorizontal_post_ci <- describe_posterior(ttest_verticalhorizontal$IEScong_vs_incon)
```

### Reliability

**Note**. Reliability analyses are conducted on all resulting indices.

#### Split-half

```{r}

compute_all_split_half <- function(df, score_vars = c("ErrorRate", "IES")) {

  conditions <- c("Congruent", "Strong", "Mild")
  # conditions <- c("Congruent", "Incongruent")
  illusions <- c("Ebbinghaus", "MullerLyer", "VerticalHorizontal")
  pairs <- list(
    c("Ebbinghaus", "MullerLyer"),
    c("VerticalHorizontal", "MullerLyer"),
    c("VerticalHorizontal", "Ebbinghaus")
  )

  # Loop through score vars and compute
  results <- list()

  for (score in score_vars) {
    cat("\n###", score, "\n")

    for (cond in conditions) {
      cat("\n##", cond, "\n")

      for (pair in pairs) {
        var1 <- paste(score, pair[1], cond, sep = "_")
        var2 <- paste(score, pair[2], cond, sep = "_")

        if (all(c(var1, var2) %in% colnames(df))) {
          cat("Pair:", var1, "vs", var2, "\n")
          res <- performance::item_split_half(df[, c(var1, var2)])
          results[[paste(score, cond, paste(pair, collapse = "_"), sep = "_")]] <- res
        } else {
          warning(paste("Missing columns:", var1, "or", var2))
        }
      }
    }
  }

  return()
}

split_half_results  <- compute_all_split_half(data_vs_wide)

```

#### Inter-illusion Reliability

Error Rate

```{r}

con_rel_er <- data_vs_wide |>
  select(matches("ErrorRate.*_Congruent")) |>
  psych::alpha()

mild_rel_er <- data_vs_wide |>
  select(matches("ErrorRate.*_mild")) |>
  psych::alpha()

strong_rel_er <- data_vs_wide |>
  select(matches("ErrorRate.*_strong")) |>
  psych::alpha()

# inc_rel_er <- data_vs_wide |>
#   select(matches("ErrorRate.*_Inongruent")) |>
#   psych::alpha()

```

```{r}

con_rel_IES <- data_vs_wide |>
  select(matches("IES.*_Congruent")) |>
  psych::alpha()

mild_rel_IES <- data_vs_wide |>
  select(matches("IES.*_mild")) |>
  psych::alpha()

strong_rel_IES <- data_vs_wide |>
  select(matches("IES.*_strong")) |>
  psych::alpha()

# inc_rel_IES <- data_vs_wide |>
#   select(matches("IES.*_Inongruent")) |>
#   psych::alpha()

```

## PCS and Illusion Sensitivity

### Error Rate

#### Ebbinghaus 
```{r}

cor_pcs_er_ebb_con <- BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = data_vs_wide$ErrorRate_Ebbinghaus_Congruent)

er_pcs_ebb_con_post <- describe_posterior(cor_pcs_er_ebb_con)

cor_pcs_er_ebb_m <- BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = data_vs_wide$ErrorRate_Ebbinghaus_Mild)

er_pcs_ebb_mild_post <- describe_posterior(cor_pcs_er_ebb_m)

cor_pcs_er_ebb_s <-  BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = data_vs_wide$ErrorRate_Ebbinghaus_Strong)

er_pcs_ebb_str_post <- describe_posterior(cor_pcs_er_ebb_s)

# Incongruent
# cor_pcs_er_ebb_inc <-  BayesFactor::correlationBF(
#   y = data_ppt$pcs,
#   x = data_vs_wide$ErrorRate_Ebbinghaus_Incongruent)
# 
# er_pcs_ebb_inc_post <- describe_posterior(cor_pcs_er_ebb_inc)

```

#### Muller-Lyer
```{r}

cor_pcs_er_mul_con <- BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = data_vs_wide$ErrorRate_MullerLyer_Congruent)

er_pcs_mul_con_post <- describe_posterior(cor_pcs_er_mul_con)

cor_pcs_er_mul_m <- BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = data_vs_wide$ErrorRate_MullerLyer_Mild)

er_pcs_mul_mild_post <- describe_posterior(cor_pcs_er_mul_m)

cor_pcs_er_mul_s <-  BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = data_vs_wide$ErrorRate_MullerLyer_Strong)

er_pcs_mul_str_post <- describe_posterior(cor_pcs_er_mul_s)

# Incongruent
# cor_pcs_er_mul_inc <-  BayesFactor::correlationBF(
#   y = data_ppt$pcs,
#   x = data_vs_wide$ErrorRate_MullerLyer_Incongruent)
# 
# er_pcs_mul_inc_post <- describe_posterior(cor_pcs_er_mul_inc)

```

#### VerticalHorizontal
```{r}

cor_pcs_er_vh_con <- BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = data_vs_wide$ErrorRate_VerticalHorizontal_Congruent)

er_pcs_vh_con_post <- describe_posterior(cor_pcs_er_vh_con)

cor_pcs_er_vh_m <- BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = data_vs_wide$ErrorRate_VerticalHorizontal_Mild)

er_pcs_vh_mild_post <- describe_posterior(cor_pcs_er_vh_m)

cor_pcs_er_vh_s <-  BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = data_vs_wide$ErrorRate_VerticalHorizontal_Strong)

er_pcs_vh_str_post <- describe_posterior(cor_pcs_er_vh_s)

# Incongruent
# cor_pcs_er_vh_inc <-  BayesFactor::correlationBF(
#   y = data_ppt$pcs,
#   x = data_vs_wide$ErrorRate_VerticalHorizontal_Incongruent)
# 
# er_pcs_vh_inc_post <- describe_posterior(cor_pcs_er_vh_inc)
```

### IES

#### Ebbinghaus 
```{r}

cor_pcs_IES_ebb_con <- BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = data_vs_wide$IES_Ebbinghaus_Congruent)

IES_pcs_ebb_con_post <- describe_posterior(cor_pcs_IES_ebb_con)

cor_pcs_IES_ebb_m <- BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = data_vs_wide$IES_Ebbinghaus_Mild)

IES_pcs_ebb_mild_post <- describe_posterior(cor_pcs_IES_ebb_m)

cor_pcs_IES_ebb_s <-  BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = data_vs_wide$IES_Ebbinghaus_Strong)

IES_pcs_ebb_str_post <- describe_posterior(cor_pcs_IES_ebb_s)

# Incongruent
# cor_pcs_IES_ebb_inc <-  BayesFactor::correlationBF(
#   y = data_ppt$pcs,
#   x = data_vs_wide$ErrorRate_Ebbinghaus_Incongruent)
# 
# IES_pcs_ebb_inc_post <- describe_posterior(cor_pcs_IES_ebb_inc)

```

#### Muller-Lyer
```{r}

cor_pcs_IES_mul_con <- BayesFactor::correlationBF(
  y = data_ppt$pcs, 
  x = data_vs_wide$IES_MullerLyer_Congruent)

IES_pcs_mul_con_post <- describe_posterior(cor_pcs_IES_mul_con)

cor_pcs_IES_mul_m <- BayesFactor::correlationBF(
  y = data_ppt$pcs, 
  x = data_vs_wide$IES_MullerLyer_Mild)

IES_pcs_mul_mild_post <- describe_posterior(cor_pcs_IES_mul_m)

cor_pcs_IES_mul_s <-  BayesFactor::correlationBF(
  y = data_ppt$pcs, 
  x = data_vs_wide$IES_MullerLyer_Strong)

IES_pcs_mul_str_post <- describe_posterior(cor_pcs_IES_mul_s)

# Incongruent
# cor_pcs_IES_mul_inc <-  BayesFactor::correlationBF(
#   y = data_ppt$pcs,
#   x = data_vs_wide$ErrorRate_MullerLyer_Incongruent)
# 
# IES_pcs_mul_inc_post <- describe_posterior(cor_pcs_IES_mul_inc)
```

#### Vertical Horizontal
```{r}

cor_pcs_IES_vh_con <- BayesFactor::correlationBF(
  y = data_ppt$pcs, 
  x = data_vs_wide$IES_VerticalHorizontal_Congruent)

IES_pcs_mul_con_post <- describe_posterior(cor_pcs_IES_vh_con)

cor_pcs_IES_vh_m <- BayesFactor::correlationBF(
  y = data_ppt$pcs, 
  x = data_vs_wide$IES_VerticalHorizontal_Mild)

IES_pcs_mul_mild_post <- describe_posterior(cor_pcs_IES_vh_m)

cor_pcs_IES_vh_s <-  BayesFactor::correlationBF(
  y = data_ppt$pcs, 
  x = data_vs_wide$IES_VerticalHorizontal_Strong)

IES_pcs_vh_str_post <- describe_posterior(cor_pcs_IES_vh_s)

# Incongruent
# cor_pcs_IES_vh_inc <-  BayesFactor::correlationBF(
#   y = data_ppt$pcs,
#   x = data_vs_wide$ErrorRate_VerticalHorizontal_Incongruent)
# 
# IES_pcs_vh_inc_post <- describe_posterior(cor_pcs_IES_vh_inc)
```

## Correlations between Illusion scores and Psychoticism

#### Ebbinghaus 
```{r}

cor_pcs_psy_ebb_con <- BayesFactor::correlationBF(
  y = data_ppt$PID5_Psychoticism,
  x = data_vs_wide$IES_Ebbinghaus_Congruent)

IES_psy_ebb_con_post <- describe_posterior(cor_pcs_psy_ebb_con)

cor_pcs_psy_ebb_m <- BayesFactor::correlationBF(
  y = data_ppt$PID5_Psychoticism,
  x = data_vs_wide$IES_Ebbinghaus_Mild)

IES_psy_ebb_mild_post <- describe_posterior(cor_pcs_psy_ebb_m)

cor_pcs_psy_ebb_s <- BayesFactor::correlationBF(
  y = data_ppt$PID5_Psychoticism,
  x = data_vs_wide$IES_Ebbinghaus_Strong)

IES_psy_ebb_str_post <- describe_posterior(cor_pcs_psy_ebb_s)

# Incongruent
# cor_pcs_psy_ebb_inc <-  BayesFactor::correlationBF(
#   y = data_ppt$pcs,
#   x = data_vs_wide$IES_Ebbinghaus_Incongruent)
# 
# IES_psy_ebb_inc_post <- describe_posterior(cor_pcs_psy_ebb_inc)
```

#### Muller-Lyer 
```{r}

cor_pcs_psy_mul_con <- BayesFactor::correlationBF(
  y = data_ppt$PID5_Psychoticism,
  x = data_vs_wide$IES_MullerLyer_Congruent)

IES_psy_mul_con_post <- describe_posterior(cor_pcs_psy_mul_con)

cor_pcs_psy_mul_m <- BayesFactor::correlationBF(
  y = data_ppt$PID5_Psychoticism,
  x = data_vs_wide$IES_MullerLyer_Mild)

IES_psy_mul_mild_post <- describe_posterior(cor_pcs_psy_mul_m)

cor_pcs_psy_mul_s <- BayesFactor::correlationBF(
  y = data_ppt$PID5_Psychoticism,
  x = data_vs_wide$IES_MullerLyer_Strong)

IES_psy_mul_str_post <- describe_posterior(cor_pcs_psy_mul_s)

# Incongruent
# cor_pcs_psy_mul_inc <-  BayesFactor::correlationBF(
#   y = data_ppt$pcs,
#   x = data_vs_wide$IES_MullerLyer_Incongruent)
# 
# IES_psy_mul_inc_post <- describe_posterior(cor_pcs_psy_mul_inc)

```

#### Vertical-Horizontal
```{r}

cor_pcs_psy_vh_con <- BayesFactor::correlationBF(
  y = data_ppt$PID5_Psychoticism,
  x = data_vs_wide$IES_VerticalHorizontal_Congruent)

IES_psy_vh_con_post <- describe_posterior(cor_pcs_psy_vh_con)

cor_pcs_psy_vh_m <- BayesFactor::correlationBF(
  y = data_ppt$PID5_Psychoticism,
  x = data_vs_wide$IES_VerticalHorizontal_Mild)

IES_psy_vh_mild_post <- describe_posterior(cor_pcs_psy_vh_m)

cor_pcs_psy_vh_s <- BayesFactor::correlationBF(
  y = data_ppt$PID5_Psychoticism,
  x = data_vs_wide$IES_VerticalHorizontal_Strong)

IES_psy_vh_str_post <- describe_posterior(cor_pcs_psy_vh_s)

# Incongruent
# cor_pcs_psy_vh_inc <-  BayesFactor::correlationBF(
#   y = data_ppt$pcs,
#   x = data_vs_wide$IES_VerticalHorizontal_Incongruent)
# 
# IES_psy_vh_inc_post <- describe_posterior(cor_pcs_psy_vh_inc)

```